\begin{enumerate}
\item
Why is security important in modern systems?
\begin{itemize}
\item 
Modern systems are typically connected to a network and are therefore potentially vulnerable to external attacks.
\end{itemize}

\item
What are some challenges of concurrency?
\begin{itemize}
\item 
Code is harder to design and to debug. Update problems may arise.
\end{itemize}

\item
What are the C++ Core Guidelines?
\begin{itemize}
\item 
Best practices that document how to build C++ systems.
\end{itemize}

\item
What's the difference between secure coding and defensive coding?
\begin{itemize}
\item 
Secure coding offers robustness to end users, whereas defensive coding offers robustness to interface consumers.
\end{itemize}

\item
How can you check whether your software contains known vulnerabilities?
\begin{itemize}
\item 
By using a CVE database or an automated scanner such as OWASP Dependency-Check or Snyk.
\end{itemize}

\item
What's the difference between static and dynamic analysis?
\begin{itemize}
\item 
Static analysis is performed on source code without executing it. Dynamic analysis requires execution.
\end{itemize}

\item
What's the difference between static and dynamic linking?
\begin{itemize}
\item 
With static linking, the executable contains all the code necessary to run the application. With dynamic linking, some parts of the code (the dynamic libraries) are shared between different executables.
\end{itemize}

\item
How you can use the compiler to fix security problems?
\begin{itemize}
\item 
Modern compilers include sanitizers that check for certain flaws.
\end{itemize}

\item
How can you implement security awareness in your CI pipeline?
\begin{itemize}
\item 
By using automated tools that scan for vulnerabilities and perform all kinds of static and dynamic analysis.
\end{itemize}
\end{enumerate}