\begin{enumerate}
\item
Why do microservices help you better use the system resources?
\begin{itemize}
\item 
It is easier to scale just the resources that are lacking instead of entire systems.
\end{itemize}

\item
How can microservices and monoliths coexist (in an evolving system)?
\begin{itemize}
\item 
New features may be developed as microservices, while some features may be split and outsourced from the monolith.
\end{itemize}

\item
Which types of teams benefit the most from microservices?
\begin{itemize}
\item 
Cross-functional autonomous teams following DevOps principles.
\end{itemize}

\item
Why is it necessary to have a mature DevOps approach when introducing microservices?
\begin{itemize}
\item 
Testing and deploying lots of microservices is almost impossible to be effected manually by separate teams.
\end{itemize}

\item
What is a unified logging layer?
\begin{itemize}
\item 
It is a configurable facility for collecting, processing, and storing logs.
\end{itemize}

\item
How do logging and tracing differ?
\begin{itemize}
\item 
Logging is usually human-readable and focused on operations, whereas tracing is usually machine-readable and focused on debugging.
\end{itemize}

\item
Why may REST not be the best choice for connecting microservices?
\begin{itemize}
\item 
It may provide bigger overhead compared to gRPC, for example.
\end{itemize}

\item
What are the deployment strategies for microservices? What are the benefits of each of them?
\begin{itemize}
\item 
Single service per host – easier to tweak the machines to the workload.

\item 
Multiple services per host – better utilization of resources.
\end{itemize}
\end{enumerate}