\begin{enumerate}
\item
What's the difference between running your applications in your cloud and making them cloud-native?
\begin{itemize}
\item 
Cloud-native design encompasses modern technologies such as containers and serverless that break the dependency on virtual machines.
\end{itemize}

\item
How can you run cloud-native applications on-premises?
\begin{itemize}
\item 
Yes, it's possible with solutions such as OpenStack, for example.
\end{itemize}

\item
What's the minimal highly available cluster size for Kubernetes?
\begin{itemize}
\item 
The minimum HA cluster requires three nodes in the control plane and three worker nodes.
\end{itemize}

\item
Which Kubernetes object represents a microservice that allows network connections?
\begin{itemize}
\item 
Service.
\end{itemize}

\item
Why is logging not sufficient in distributed systems?
\begin{itemize}
\item 
Gathering logs and looking for correlations between them in distributed systems is problematic. Distributed tracing is better suited for certain use cases.
\end{itemize}

\item
How does a service mesh help with building secure systems?
\begin{itemize}
\item 
A service mesh abstracts connectivity between different systems, which allows encryption and auditing to be applied.
\end{itemize}

\item
How does GitOps increase productivity?
\begin{itemize}
\item 
It uses a familiar tool, Git, to handle the CI/CD without the need to write dedicated pipelines.
\end{itemize}

\item
What's the standard CNCF project for monitoring?
\begin{itemize}
\item 
Prometheus.
\end{itemize}
\end{enumerate}