\begin{enumerate}
\item
What are quality attributes?

\begin{itemize}
\item 
A mediator "mediates" between services, so it needs to know how to process each request. A broker only knows where to send each request, so it's a lightweight component. It can be used to create a publishersubscriber (pub-sub) architecture.
\end{itemize}

\item
What sources should be used when gathering requirements?

\begin{itemize}
\item 
The context of your system, existing documentation, and the system's stakeholders.
\end{itemize}

\item
How can you tell whether a requirement is architecturally significant?

\begin{itemize}
\item 
Architecturally significant requirements (ASRs) often require a separate software component, impact a large part of the system, are hard to achieve, and/or force you to make trade-offs.
\end{itemize}

\item
How should you document graphically the functional requirements various parties may have regarding your system?

\begin{itemize}
\item 
Prepare a use case diagram.
\end{itemize}

\item
When is development view documentation useful?

\begin{itemize}
\item 
In cases where you're developing a large system with many modules and need to communicate global constraints and common design choices to all the software teams.
\end{itemize}

\item
How should you automatically check whether your code's API documentation is out of date?

\begin{itemize}
\item 
Doxygen has built-in checks, like the one that warns you about mismatches between the function signatures and their parameters in comments
\end{itemize}

\item
How should you show on a diagram that a given operation is handled by different components of the system?

\begin{itemize}
\item 
Use one of the UML interaction diagrams for this purpose. Sequence diagrams are a good choice, although communication diagrams can be fine in certain scenarios, too.
\end{itemize}

\end{enumerate}