\begin{enumerate}
\item
What is event sourcing?
\begin{itemize}
\item 
This is an architectural pattern that relies on keeping track of events that change the state of the system instead of keeping track of the state per se. It brings benefits such as lower latency, free audit logs, and debugability.
\end{itemize}

\item 
What are the practical consequences of the CAP theorem?
\begin{itemize}
\item 
As network partitions happen, if you want a distributed system, you'll need to choose between consistency and availability. In cases of partitions, you can either return stale data, an error, or risk timeouts.
\end{itemize}

\item 
What can you use Netflix's Chaos Monkey for?
\begin{itemize}
\item 
It can help you prepare for unexpected downtime of your services.
\end{itemize}

\item 
Where can caching be applied?
\begin{itemize}
\item 
Either on your client's side, in front of web servers, databases, or applications, or on a host near your potential client, depending on your needs.
\end{itemize}

\item 
How do you prevent your app from going down when a whole data center does?

\begin{itemize}
\item 
By using geodes.
\end{itemize}

\item 
Why use an API Gateway?
\begin{itemize}
\item 
To simplify client code, as it doesn't need to hardcode the addresses of your service instances.
\end{itemize}

\item How can Envoy help you to achieve various architectural goals?	
\begin{itemize}
\item 
It aids your system's fault tolerance by providing backpressure, circuit breaking, automatic retries, and outlier detection.

\item 
It aids deployability by allowing canary releases and blue-green deployments.

\item 
It also offers load balancing, tracing, monitoring, and metrics.
\end{itemize}
\end{enumerate}