\begin{enumerate}
\item
What's the difference between installing and exporting your targets in CMake?
\begin{itemize}
\item 
Exporting means the targets will be available for other projects that try to find our package, even if our code is not installed. CMake's package registry can be used to store data about locations of the exported targets. The binaries never leave the build directory. Installation requires the targets to be copied somewhere and, if it's not a system directory, setting up paths to the config files or the targets themselves.
\end{itemize}

\item
How do you make your template code compile faster?
\begin{itemize}
\item 
Follow the Rule of Chiel.
\end{itemize}

\item
How do you use multiple compilers with Conan?
\begin{itemize}
\item 
Use Conan profiles.
\end{itemize}

\item
What do you do if you'd like to compile your Conan dependencies with the preC++11 GCC ABI?
\begin{itemize}
\item 
Set compiler.libcxx to libstdc++ instead of libstdc++11.
\end{itemize}

\item
How to ensure we force a specific C++ standard in CMake?
\begin{itemize}
\item 
By calling set\_target\_properties(our\_target PROPERTIES
CXX\_STANDARD our\_required\_cxx\_standard
CXX\_STANDARD\_REQUIRED YES CXX\_EXTENSIONS NO).
\end{itemize}

\item
How would you build documentation in CMake and ship it along with your RPM package?
\begin{itemize}
\item 
Create a target to generate the documentation as described in Chapter 3, Functional and Nonfunctional Requirements, install it to CMAKE\_INSTALL\_DOCDIR, and then make sure the path is not specified in the CPACK\_RPM\_EXCLUDE\_FROM\_AUTO\_FILELIST variable.
\end{itemize}
\end{enumerate}